Does at-issueness predict projection? Further investigations of the Gradient Projection Principle

### 2022-07-06 writing/investigating-GPP.tex
update paper

@@

@@

Previous investigations of the relationship between projection and at-issueness:

@@

@@

\item p.6: This test brings up the question of what counts as a direct response. One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter & Asher, 2016).

@@

\item p.7: A third notion of at-issueness is situated within theories which view discourse as structured by coherence relations, also called ?rhetorical? or ?discourse? relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann & Thompson, 1988).

@@

\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia & McConnell-Ginet, 2000; Karttunen & Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel & Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett & Koev, 2015; Tonhauser, 2012).

@@

\item \citealt{snider-thesis, snider2017, snider2018}

>>>>

\begin{exe}
\ex Context: A and B live near a dog that regularly barks for extended periods of time. It is barking right now. A says:  \\ Ugh. The damn dog is barking again.
\end{exe}

Contents:

\begin{itemize}
\item there is a unique salient dog (backgrounded, ps)
\item the dog has barked before (backgrounded, ps)
\item the speaker is annoyed at the dog/barking (new, could be what the speaker wants to contribute)
\item the dog is barking (``at-issue'', ordinary entailment)
\end{itemize}

First dimension: Role in discourse context: some anchor to context, some convey relevant backgrounded information, some are what the speaker wants to communicate.

Second dimension: Compositional semantics: some interact, some don't.

Third dimension: Role in subsequent discourse: some are what carries the conversation on, some do not.

One of the contents is privileged in that it is the content that is put up for negotiation. It is often new information, though not necessarily. It is what is ``linguistically relevant'': it interacts with operators and is accessible for anaphoric reference.

@@

We already know that the GPP is stated too strongly: prior influences projection (\citealt{degen-tonhauser-openmind}) {\bf independently of at-issueness (unpublished)}. At-issueness is one factor among many.

@@

Previous investigations of/claims about the relationship between projection and at-issueness:

@@

- Esipova, Maria. 2018. QUD-addressing appositives don?t have to be clause-final. Snippets 33. 7?8. https://doi.org/10.7358/snip-2018-033-esip.

- Djärv, Kajsa. 2019. Factive and assertive attitude reports. Philadelphia, PA: Uni- versity of Pennsylvania dissertation. https://repository.upenn.edu/edissertations/ 3645.

@@

\item p.6: This test brings up the question of what counts as a direct response. One peculiarity about the responses in 10?11 is that they are anaphoric to parts of the target utterance through propositional anaphors like that or no.7 This is important because nonanaphoric denials behave differently: They can felicitously target nominal appositives (see also Hunter \& Asher, 2016).

@@

\item p.7: A third notion of at-issueness is situated within theories which view discourse as structured by coherence relations, also called ?rhetorical? or ?discourse? relations (Asher \& Lascarides, 2003; Hobbs, 1979, 1985; Kehler, 2002; Mann \& Thompson, 1988).

@@

\item endnote 5: The assent/dissent test is known under different names in the literature and has been used to distinguish asserted content from implications triggered by constructions as different as presupposition markers (Chierchia \& McConnell-Ginet, 2000; Karttunen \& Peters, 1979; Shanon, 1976; Strawson, 1950), epistemic modals (Lyons, 1977; Papafragou, 2006; von Fintel \& Gillies, 2007), evidentials (Faller, 2002; Matthewson et al., 2007; Murray, 2014), and appositives (AnderBois et al., 2015; Koev, 2013; Syrett \& Koev, 2015; Tonhauser, 2012).

@@

\item \citealt{snider-thesis}

\item \citealt{snider2017} (superseded for the most part by 2018 paper)

\item In this paper, I argue that there is no tight link between at-issueness and anaphoric availability? that these two properties of propositions are distinct. I present data which demonstrates that a proposition?s at-issue status alone is not sufficient to determine whether it will be available for anaphoric reference (including but not limited to direct assent/dissent), as well as data which shows that a proposition?s being available for anaphoric reference cannot be used to diagnose its at-issue status.

\item  Tonhauser asserts that they are all indeed testing for at-issueness: ?they are all useful to diagnose (not-)at-issueness with at least one kind of content? (p. 252). {\bf I was wrong}

\item p.6: The question/answer tests return the same results, because their use of anaphors (response particles) is only incidental. In contrast, the direct dissent tests rely on explicit anaphors by design; an anaphor-less equivalent to the direct dissent test in (11) ceases to be the same diagnostic.

\item \citealt{snider2018}

\item p.374: at-issueness and the anaphoric potential of a proposition are distinct notions; one class of tests commonly used to diagnose at-issue status in fact diagnoses only anaphoric potential

\item NRRCs: anaphoric potential varies, especially with position, but never good to address question under discussion

\item at-issue content: certain content conveyed by an utterance has a distinguished status as the main point of the utterance, while other content does not. Potts 2005 discusses ?at-issue entailments? as distinct from presuppositions and conventional implicatures, but does not describe how to identify content as at-issue.

\item p.375: Properties of at-issue content (from \citealt{tonhauser-sula6})

At-issue content can be directly assented or dissented with {\bf third dimension}

At-issue content addresses the question under discussion {\bf first dimension}

At-issue content determines the relevant set of alternatives {\bf second dimension, compositionality}

Some content may overlap on all three dimensions: e.g., that Fritz is happy in {\em Fritz is happy}. But sentence-final appositives can be assented/dissented with, but they do not address the QUD or determine the relevant set of alternatives. And that the speaker is annoyed may address the QUD, but does not determine the relevant set of alternatives.

\item p.376f: Snider argues against \citealt{syrett-koev2015} claim that appositives can address the QUD. ``Why questions are crucially different from these other questions in that they seek explanations. Because they seek explanations, such questions have no single comprehensive answer: any answer given is defeasible2 or can be considered insufficiently precise, and additional contextual information can always contribute to an explanation (see discussion in Bjorndahl \& Snider, 2016). The appositive might contribute information which bolsters an explanation, supporting the matrix clause?s answer to the QUD, but that alone does not constitute addressing the QUD. If the matrix clause does not address the QUD, then an appositive alone cannot be taken to address the QUD, even in a why question, as in (7)''

Syrett \& Koev 2015:587 further points to an example where an appositive appears to ?address one part of a coordinated QUD?, inspired by an example from Koev 2013. But the concept of a ?coordinated QUD? is novel: Roberts 1996 describes no such structure, instead describing sequences of ?relevant sub-questions to some super-question? as an ?enumeration, suggesting a plan for how to attack the super-question?, and says that one ?can only address such sub-questions one at a time? (5). The response in (8A) thus addresses the first sub-question?as it must, because the second requires an answer to the first?and not both at once.

\item p.378: The crucial difference between the classes of diagnostics illustrated above is that the first is anaphora-based, where the other two are QUD-based, as they test what can address or establish a Question Under Discussion.

Snider assumes that the two QUD diagnostics (addressing QUD, establishing QUD) are QUD-based, i.e., tap into the same property? But establishing a QUD is about compositionality, and addressing a QUD is about the role of content in discourse.

\item p.380: Alternatively, this observation might lead one to move away from the Simons et al. 2010 definition entirely. Hunter \& Asher 2016 proposes a discourse based account using SDRT (Asher, 1993; Asher \& Lascarides, 2003) in which at-issueness and anaphoric potential are indeed closely linked?which isn?t at odds with the data discussed here, because they use an entirely different notion of at-issueness. On this account, ?content is at-issue at a certain point of the discourse just in case it is on the [Right Frontier (RF)] of the discourse at that time. Content that is AI becomes NAI when knocked off the RF; content that is NAI can become AI if targeted through discourse subordination? and thus returned to the Right Frontier (Hunter \& Asher, 2016:1036). And, because discourse referents (propositional and otherwise) are only accessible for anaphora resolution in SDRT if they are on the Right Frontier, propositional content which is available for anaphoric reference is necessarily at-issue on this account.

\item p.380: However one chooses to define or model at-issueness, it is important to recognize the difference between the at-issue status of a proposition and its anaphoric potential. The two are at least conceptually distinct: at-issueness is about what the main point of an utterance is, while anaphoric potential is about what propositions are available for subsequent anaphoric reference.
